\chapter{Introduction}
\label{ch:intro}

\section{Overview}
This report focuses on the biotechnological production of \textbf{1,3-propanediol (1,3-PDO)} using a genetically engineered \textit{Escherichia coli} system. 
1,3-PDO (\ce{C3H8O2}) is a three-carbon diol that serves as a versatile building block in polymer chemistry, particularly in the synthesis of polytrimethylene terephthalate (PTT)\cite{nimbalkarReviewMicrobial12024}. 
Its attractive properties, such as low toxicity, high miscibility with water and alcohols, and suitability for polycondensation reactions, make it an important industrial chemical.

\section{Motivation}
\subsection{Sustainability}
The biotechnological route to PDO production is attractive due to its renewable feedstocks and lower carbon footprint compared to petrochemical synthesis routes from acrolein\cite{TJAHJASARI2011258} or using ethylene oxide as done by Shell\cite{TJAHJASARI2011258}. 
This supports global sustainability goals and the transition toward green chemistry.

\subsection{Industry}
PDO serves as a monomer for the bio-based polymer PTT (commercially known as \textit{Sorona\textsuperscript{\textregistered}}), used in textile fibers, coatings, and engineering plastics.
Its market is growing steadily as industries move towards sustainable materials \cite{ewingFermentationProductionBiobased2022}.

\section{Scope of Study}
In this report, I have tried to study and understand several important aspects of chemical and biochemical engineering, such as:
\begin{itemize}
    \item \textbf{Kinetic modeling} to analyze fermentation reaction behavior.
    % \item \textbf{Mass and energy balances} across different unit operations involved in the process.
    \item \textbf{Reactor design} considerations for biological systems under aerobic and anaerobic conditions.
    \item \textbf{Downstream separation and purification} techniques used to recover and refine the final product.
\end{itemize}
The overall aim was to study the available the biological pathways involved in production with process design and industrial feasibility.

